Systematically Optimizing Abstract Abstract Machines

### 2013-03-29 paper/paper.tex
Some changes to intro.

\begin{abstract}
Abstracting abstract machines has been proposed as a
lightweight approach to designing sound and computable program
analyses.  The approach derives abstract interpreters from
existing machine semantics and has been applied to a variety of
languages with features widely considered difficult to analyze.
Although sound analyzers are straightforward to build under this approach,
they are also prohibitively inefficient.

This article contributes a step-by-step process for going from a
\naive{} analyzer derived under the abstracting abstract machine
approach to an efficient program analyzer.  The end result of the
process is a two to three order-of-magnitude improvement over the
systematically derived analyzer, making it competitive with
hand-optimized implementations that compute fundamentally less precise
results.
\end{abstract}

@@

The \emph{abstracting abstract machines} (AAM)
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides a systematic way of transforming a
programming language semantics in the form of an abstract machine into a family
of abstract interpreters.
%
The approach parameterizes these families with policies for regulating analytic
precision.
%
While flexible and robust, the AAM approach unfortunately yields analyzers with
poor performance relative to hand-optimized analyzers.
%
Our work takes aim squarely at this ``efficiency gap,'' and narrows it in an
equally systematic way.

@@

However, there is a considerable weakness to the approach: an
analyzer designed and implemented by following the AAM recipe is
prohibitively inefficient without both further approximation and
further implementation effort.

In this article, we develop a systematic approach to deriving a
practical implementation of an abstract-machine-based analyzer using
mostly semantic means - not tricky engineering. Our goal is to empower
programming language implementers and researchers to explore and
convincingly exhibit their ideas with a low barrier to entry. The
optimizations we describe are widely applicable and apparently
effective to scale far beyond toy examples.

>>>>

\begin{abstract}
The technique of \emph{abstracting abstract machines} (AAM) provides a
systematic approach for deriving computable approximations of
evaluators that are easily proved sound.
%
This article contributes a complementary step-by-step process for
subsequently going from a \naive{} analyzer derived under the AAM
approach, to an efficient and correct implementation.  The end result
of the process is a two to three order-of-magnitude improvement over
the systematically derived analyzer, making it competitive with
hand-optimized implementations that compute fundamentally less precise
results.
\end{abstract}

@@

Program analysis provides sound predictive models of program behavior, but
in order for such models to be effective, they must be efficiently
computable and correct.  Past approaches to designing program analysis
have often featured abstractions that are far removed from the
original language semantics, requiring ingenuity in their construction
and effort in their verification.
%
The \emph{abstracting abstract machines} (AAM)
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides an alternative: a systematic way
of transforming a programming language semantics in the form of an
abstract machine into a family of abstract interpreters.  It thus reduces
the burden of constructing and verifying the soundness of an abstract
interpreter.
%

@@

Unfortunately, the AAM approach yields analyzers with poor performance
relative to hand-optimized analyzers.
%
Our work takes aim squarely at this ``efficiency gap,'' and narrows it
in an equally systematic way through a number of simple steps, many of
which are inspired by run-time implementation techniques such as
laziness and compilation to avoid interpretative overhead.  Each of
these steps is proven correct, so the end result is an implementation
that is trustworthy and efficient.

In this article, we develop a systematic approach to deriving a
practical implementation of an abstract-machine-based analyzer using
mostly semantic means rather than tricky engineering. Our goal is to
empower programming language implementers and researchers to explore
and convincingly exhibit their ideas with a low barrier to entry. The
optimizations we describe are widely applicable and apparently
effective to scale far beyond the size of programs typically considered
in the recent literature on flow analysis for functional languages.
